Dish: Distribution-oblivious Shell Scripting

### 2020-02-21 pldi/draft.tex
Notes on section 4

\begin{itemize}
\item Why is parallelism important (especially in the context of data
  processing). Significant performance benefits as most computer
  systems today have multiple processors that are under utilized
\item (Similarly to how it is now) unfortunately, it is quite hard
  reaping these benefits in contrast to sequential data
  processing. [Show Bash pipeline].
\item To benefit from parallelism, the current options are
  extreme. Either very fine-grained and low level (parallelism using
  fine-grained C libraries) for maximum performance, or very coarse
  grained and bulky where one has to use a proper big-data framework.
\item The common thing is that one needs to rewrite big part of their
  application, often requiring expertise and redesigning its
  architecture, in order to get benefits that in the end might not
  explain the cost.
\item To address this issue, we propose Dish, a thin layer that
  automatically! parallelizes a given shell script. Key insight is
  that a large subset of the shell language can be encoded in a
  dataflow graph model, a flexible representation that allows for
  parallelization optimizations.
\item Study of the shell primitives, identification of
  families/classes of commands that share
  distributability/parallelization properties, and a classification of
  coreutils/POSIX commands in these distributability properties.
\item Dataflow graph model that is suitable for the shell
\item A set of optimizations on this model that can be thought of as
  rewriting rules on the dataflow graph, that are sound and exploit
  parallelism.
\end{itemize}

@@

The core of \sys is a dataflow graph model, inspired by the ones used in popular distributed stream processing systems~\sx{related}.
% such as MapReduce \cite{mapreduce:08} and Apache Spark \cite{spark:12}).
These systems introduce data parallelism by parallelizing nodes of the dataflow graph.
Unfortunately, % their optimizations in many of these systems 
 they often do not preserve the model's semantics, blurring the lines between specification, optimization, and implementation.
In contrast, \sys's dataflow model inherently supports the data parallelism found in shell pipelines.
More specifically, it develops a set of parallelization-exposing optimizations represented as semantics-preserving graph transformations, effectively exposing data parallelism as task parallelism.
The \sys optimizer then applies these transformations in an iterative way, optimizing the program to exploit the available computational resources.

%% The standard dataflow graph exposes task parallelism. Different nodes
%% in the graph can be executed concurrently, communicating through the
%% edges.
%% Task level parallelism is already exposed by the shell, by using
%% pipes \ttt{|} and the background operator \ttt{\&}.
%% Unfortunately, simple task parallelism leaves a lot to be desired,
%% since scaling a computation requires manually introducing more
%% operators.

%% In order to adequately utilize the available computational resources,
%% a distributed implementation should exploit data parallelism. Data
%% parallelism is based on the observation that some operations produce
%% the same result when performed independently on subsets of the input
%% data, merging the results afterwards. Exploiting data parallelism is
%% one of the major reasons that popular distributed stream processing
%% systems (such as MapReduce \cite{mapreduce:08} and Apache Spark
%% \cite{spark:12}) achieve significant performance.

%% A very important consequence of this is that we can iteratively apply
%% optimizations that expose data parallelism, or even revert them when
%% the resulting dataflow graph is too wide. This flexibility unifies the
%% unoptimized and optimized versions of a computation, enabling thinking
%% about them with the same conceptual tools.

In order to effectively apply optimizations and distribute a shell script, \sys first translates it to a manipulable representation.
\sys first introduces a variant of the standard dataflow graph model as a candidate intermediate representation~\sx{graph-components}.
The main benefit of this model is that it exposes both task and data parallelism.
It then defines a set of optimizations that expose and exploit the data parallelism of commands in \sta and \pur as semantics-preserving graph transformations~\sx{ir:transformations}.

@@

A node $f$ of the graph represents a function from one (possibly empty)
input stream to an output stream $f : D\kstar \to D\kstar$, where $D$
represents the basic data type of a line of characters. This
representation captures the majority of shell commands. We require that
the function $f$ is monotone w.r.t.\ the prefix order. This captures
the idea that a node cannot retract output after it has been emitted.
%More formally, $\forall f, x, y, f(x) = y$, then for any $x'$, there exists $y'$ such that $f(x.x') = y.y'$, where $.$ is standard
%concatenation and will be omitted when obvious.
An example of a node is the following command:

@@

@@

@@

\paragraph{Annotation-based Distribution}
Recent systems on light-touch distribution, such as Ignis~\cite{ignis:19} and Mozart~\cite{mozart:19}, are more general than cluster-computing frameworks.
They also lower (but not entirely eliminate) developer effort by automating most of the rewriting process.
% Both work on high-level, managed languages---server-side JavaScript for Ignis and Python for Mozart---that allow runtime transformations without affecting the broader environment under which a program executes.
For the critical points where user input is required, they provide a high-level declarative DSL with varying (but limited) degrees of expressive power.
% Annotations are used by ---complicated by the use of third-party libraries similar to \sys's extensions.
Unfortunately, a challenge with this approach is that users composing applications by using third-party libraries do not understand their distributability details well-enough to annotate them.

@@

>>>>

\begin{itemize}
\item Why is parallelism important (especially in the context of data
  processing). Significant performance benefits as most computer
  systems today have multiple processors that are under utilized
\item (Similarly to how it is now) unfortunately, it is quite hard
  reaping these benefits in contrast to sequential data
  processing. [Show Bash pipeline].
\item To benefit from parallelism, the current options are
  extreme. Either very fine-grained and low level (parallelism using
  fine-grained C libraries) for maximum performance, or very coarse
  grained and bulky where one has to use a proper big-data
  framework. (Can we fit the shell ones here (gnu-parallel, etc))
\item The common thing is that one needs to rewrite big part of their
  application, often requiring expertise and redesigning its
  architecture, in order to get benefits that in the end might not
  explain the cost.
\item To address this issue, we propose Dish, a thin layer that
  automatically! parallelizes a given shell script. Key insight is
  that a large subset of the shell language can be encoded in a
  dataflow graph model, a flexible representation that allows for
  parallelization optimizations.
\item Contributions:
\item Study of the shell primitives, identification of
  families/classes of commands that share
  distributability/parallelization properties, and a classification of
  coreutils/POSIX commands in these distributability properties.
\item Dataflow graph model that is suitable for the shell
\item A set of optimizations on this model that can be thought of as
  rewriting rules on the dataflow graph, that are sound and exploit
  parallelism.
\end{itemize}

@@

The core of \sys is a dataflow graph model, inspired by the ones used
in popular distributed stream processing systems~\sx{related}. The
dataflow model can express a large subset of Shell programs and
inherently supports the data parallelism found in shell pipelines by
exposing it as task parallelism. More preciely, we have developed a
set of parallelization-exposing optimizations represented as
semantics-preserving graph transformations. These transformations can
then be applied in an iterative way, optimizing the program to utilize
available computational resources.

%% In order to effectively apply optimizations and distribute a shell
%% script, \sys first translates it to a manipulable representation.

\kk{It might be beneficial to do a full pass over this section
  describing the new model (where a node that has many sequential
  inputs is always prepended with a cat node.}

@@

A node $f$ of the graph represents a function from one (possibly
empty) input stream to an output stream $f : D\kstar \to D\kstar$,
where $D$ represents the basic data type of a line of characters
\kk{We have to extend this to more general nodes. It has to be from
  many inputs to many outputs. The commands that we optimize at the
  moment are the ones that can be rewritten to take one stream input
  to one stream output, and then some other inputs and outputs that
  are considered static.}. This representation captures the majority
of shell commands. We require that the function $f$ is monotone
w.r.t.\ the prefix order. This captures the idea that a node cannot
retract output after it has been emitted.
%More formally, $\forall f, x, y, f(x) = y$, then for any $x'$, there
%exists $y'$ such that $f(x.x') = y.y'$, where $.$ is standard
%concatenation and will be omitted when obvious.
An example of a node is the following command:

@@

\kk{I think that we should probably add the ``new'' split optimization
  (and maybe any others that we develop until then) here.}

@@

\kk{Ideally we would want to talk here about the as-late-as possible
  dish translation that allows us to be safe despite the Shell's
  dynamic nature. I think that this is a great point about the
  work. Maybe we should also focus on implementing it before the OSDI
  deadline.}

@@

\paragraph{Annotation-based Distribution}
Recent systems on light-touch distribution, such as
Ignis~\cite{ignis:19} and Mozart~\cite{mozart:19}, are more general
than cluster-computing frameworks.  They also lower (but not entirely
eliminate) developer effort by automating most of the rewriting
process.
% Both work on high-level, managed languages---server-side JavaScript for Ignis and Python for Mozart---that allow runtime transformations without affecting the broader environment under which a program executes.
For the critical points where user input is required, they provide a
high-level declarative DSL with varying (but limited) degrees of
expressive power.
% Annotations are used by ---complicated by the use of third-party
% libraries similar to \sys's extensions.
Unfortunately, a challenge with this approach is that users composing
applications by using third-party libraries do not understand their
distributability details well-enough to annotate them.

@@

%% Possible points to add: These systems introduce data parallelism by
%% parallelizing nodes of the dataflow graph.  Unfortunately, they
%% often do not preserve the model's semantics, blurring the lines
%% between specification, optimization, and implementation.  In
%% contrast, \sys's dataflow model inherently supports the data
%% parallelism found in shell pipelines.  More specifically, it
%% develops a set of parallelization-exposing optimizations
%% represented as semantics-preserving graph transformations,
%% effectively exposing data parallelism as task parallelism.
